---
title: "compare_with_gsa"
author: "Takeo Naito, MD PhD"
date: "8/10/2019"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(data.table)
library(readr)
library(knitr)
library(kableExtra)
```
# prepare adimix file and target file(include perianal information)

```{r}
admix <- fread("/mnt/share6/FOR_Takeo/array_139/admix_with_jewish_nonjewish_prop_added.txt")

target <- fread("/mnt/share6/FOR_Takeo/WES/SKAT/SKAT_target_rev2.txt") %>% 
  select(IID,GeneticID,anal)
target$GeneticID <- str_replace(target$GeneticID,"-","0")
```
# read phenotype file which includes Race information (self reported)

```{r}
pheno <- read_xls("../data/Copy of Genetics 01_02_2019.xls",col_types = "text")
colnames(pheno) <- make.names(colnames(pheno))
pheno <- pheno %>% 
  distinct(Genetic.ID, .keep_all = TRUE) %>% 
  select(Genetic.ID, Race)
pheno$Genetic.ID <- str_replace(pheno$Genetic.ID,"-","0")
```

```{r}
target <- target %>% 
  left_join(pheno,by=c("GeneticID" = "Genetic.ID"))

```
# merge each file and extract only Caucasian (self reported)

```{r}
target1 <- target %>% 
  inner_join(admix,by = c("GeneticID" = "FID")) %>% 
  filter(anal != -9) %>% 
  filter(Race == "Caucasian")
```

# Some outlier exist

```{r}
ggplot(target1,aes(Caucasian)) + geom_histogram()
```



# filter by admix Caucasian
```{r}
target2 <- target1 %>% 
  filter(Caucasian >= 0.75)
```
# Check outlier's (Caucasian by self reported but not Caucasian by admixture )
```{r}
outlier <- target1 %>% 
  filter(Caucasian < 0.75)

dim(outlier)
outlier
```

# select only Caucasian by admixture

```{r}
broad <- fread("/mnt/share6/FOR_Takeo/GSA/cedars_broad_GSA_KEY copy.txt")
feinstein <- fread("/mnt/share6/FOR_Takeo/GSA/cedars_feinstein_GSA_KEY copy.txt")
```
```{r}
broad1 <- broad %>% 
  mutate(id = paste0(broad$geneticID,broad$geneticID_viaNIDDK)) %>% 
  select(vcf_ID,id) %>% 
  distinct(id,.keep_all = TRUE)

broad1$id <- str_replace(broad1$id,"-","0")
```
```{r}
feinstein$geneticID <- str_replace(feinstein$geneticID,"-","0")
feinstein <- feinstein %>% 
  distinct(geneticID,.keep_all = TRUE)
```

```{r}
target3 <- target2 %>% 
  left_join(broad1,by = c("GeneticID" = "id")) %>% 
  left_join(feinstein,by =c("GeneticID" = "geneticID"))

target4 <- target3 %>% 
  drop_na(vcf_ID.y)
```


```{r}
target4 %>% 
  select(IID.x) %>% 
  write_tsv("/mnt/share6/FOR_Takeo/WES/perianal/caucasian_common.txt",col_names = FALSE)


target4 %>% 
  select(vcf_ID.y) %>% 
  write_tsv("/mnt/share6/FOR_Takeo/GSA/caucasian_common.txt",col_names = FALSE)
```



```{bash}
plink2 \
--bfile ../data/cleaned_cedars_bialle \
--exclude range ../data/highLDregions.txt \
--geno 0.01 \
--indep-pairwise 50 5 0.1 \
--keep ../data/caucasian_admix.txt \
--maf 0.05 \
--mind 0.01 \
--not-chr X Y \
--out ../filtering/hail/cleaned_cedars_cauc_rev

```


```{bash}
plink2 \
--bfile ../data/cleaned_cedars_bialle \
--extract ../filtering/hail/cleaned_cedars_cauc_rev.prune.in \
--keep ../data/caucasian_admix.txt \
--out ../filtering/hail/PCAclean_cauc_rev \
--pca 5 \
--threads 10
```



```{bash}
plink2 \
--covar ../filtering/hail/PCAclean_cauc_rev.eigenvec \
--covar-col-nums 3,4,5,6 \
--logistic hide-covar \
--out /mnt/share6/FOR_Takeo/WES/perianal/anal_common \
--bfile ../data/cleaned_cedars \
--pheno /mnt/share6/FOR_Takeo/WES/SKAT/SKAT_target_rev2.txt \
--pheno-col-nums 29 \
--keep /mnt/share6/FOR_Takeo/WES/perianal/caucasian_common.txt \
--adjust \
--threads 4
```
```{bash}
plink2 \
--keep-if anal = 2 \
--freq \
--out /mnt/share6/FOR_Takeo/WES/perianal/case_common \
--bfile ../data/cleaned_cedars \
--keep /mnt/share6/FOR_Takeo/WES/perianal/caucasian_common.txt \
--pheno /mnt/share6/FOR_Takeo/WES/SKAT/SKAT_target_rev2.txt 
```


```{bash}
plink2 \
--keep-if anal = 1 \
--freq \
--out /mnt/share6/FOR_Takeo/WES/perianal/cnt_common \
--bfile ../data/cleaned_cedars \
--keep /mnt/share6/FOR_Takeo/WES/perianal/caucasian_common.txt \
--pheno /mnt/share6/FOR_Takeo/WES/SKAT/SKAT_target_rev2.txt 
```

```{r}
res <- fread("/mnt/share6/FOR_Takeo/WES/perianal/anal_common.anal.glm.logistic") %>% 
  arrange(P)

casefre <- fread("/mnt/share6/FOR_Takeo/WES/perianal/case_common.afreq")
colnames(casefre) <- make.names(colnames(casefre))

cntfre <- fread("/mnt/share6/FOR_Takeo/WES/perianal/cnt_common.afreq")
colnames(cntfre) <- make.names(colnames(cntfre))

```


```{r}
merge1 <- casefre %>% 
  inner_join(cntfre,by = "ID") %>% 
  select(-X.CHROM.y,-REF.y,-ALT.y)


merge2 <- res %>% 
  inner_join(merge1,by = "ID") %>% 
  select(X.CHROM.x, POS, REF,ALT,OBS_CT,OR,P,ALT_FREQS.x,OBS_CT.x,
         ALT_FREQS.y,OBS_CT.y) %>%  
  dplyr::rename(CHR =  X.CHROM.x, freq_case = ALT_FREQS.x, freq_cnt = ALT_FREQS.y,
         total_n = OBS_CT, case_n = OBS_CT.x, cnt_n = OBS_CT.y) %>% 
  mutate(case_n = case_n/2, cnt_n = cnt_n/2)
```


```{r}
kable(head(merge2,10)) %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```

```{r}
head(merge2)
```



```{r}
target4 %>% 
  mutate(FID = 0) %>% 
  select(FID,vcf_ID.y,anal) %>% 
  rename(IID =vcf_ID.y ) %>% 
  write_tsv("/mnt/share6/FOR_Takeo/GSA/ferrintei/caucasian_common_pheno.txt")
```


```{bash}
plink2 \
--keep-if anal = 2 \
--freq \
--snp 20:55025762 \
--out /mnt/share6/FOR_Takeo/GSA/ferrintei/case_common \
--pfile /mnt/share6/FOR_Takeo/GSA/ferrintei/test_feinstein \
--keep /mnt/share6/FOR_Takeo/GSA/caucasian_common.txt \
--pheno /mnt/share6/FOR_Takeo/GSA/ferrintei/caucasian_common_pheno.txt
```


```{bash}
plink2 \
--keep-if anal = 1 \
--freq \
--snp 20:55025762 \
--out /mnt/share6/FOR_Takeo/GSA/ferrintei/cnt_common \
--pfile /mnt/share6/FOR_Takeo/GSA/ferrintei/test_feinstein \
--keep /mnt/share6/FOR_Takeo/GSA/caucasian_common.txt \
--pheno /mnt/share6/FOR_Takeo/GSA/ferrintei/caucasian_common_pheno.txt
```

# confirm frequency consistency in GSA

```{r}
case_ferri <- fread("/mnt/share6/FOR_Takeo/GSA/ferrintei/case_common.afreq")
cnt_ferri <- fread("/mnt/share6/FOR_Takeo/GSA/ferrintei/cnt_common.afreq")

case_ferri
cnt_ferri

  
```


```{bash}
plink2 \
--keep-if anal = 1 \
--export A \
--snp 20:55025762 \
--out /mnt/share6/FOR_Takeo/GSA/ferrintei/rs6069753_cnt \
--pfile /mnt/share6/FOR_Takeo/GSA/ferrintei/test_feinstein \
--keep /mnt/share6/FOR_Takeo/GSA/caucasian_common.txt \
--pheno /mnt/share6/FOR_Takeo/GSA/ferrintei/caucasian_common_pheno.txt
```

```{bash}
plink2 \
--keep-if anal = 2 \
--export A \
--snp 20:55025762 \
--out /mnt/share6/FOR_Takeo/GSA/ferrintei/rs6069753_case \
--pfile /mnt/share6/FOR_Takeo/GSA/ferrintei/test_feinstein \
--keep /mnt/share6/FOR_Takeo/GSA/caucasian_common.txt \
--pheno /mnt/share6/FOR_Takeo/GSA/ferrintei/caucasian_common_pheno.txt
```


```{bash}
plink2 \
--export A \
--snp 20:55025762:G:A \
--out ~/WES/result/rs6069753_cnt \
--bfile ~/WES/data/cleaned_cedars \
--keep ~/WES/data/caucasian_admix.txt \
--pheno /mnt/share6/FOR_Takeo/WES/SKAT/SKAT_target_rev2.txt \
--pheno-col-nums 29 
```




```{r}
gsa <- fread("/mnt/share6/FOR_Takeo/GSA/ferrintei/rs6069753_cnt.raw")
colnames(gsa) <- make.names(colnames(gsa))

gsa1 <- fread("/mnt/share6/FOR_Takeo/GSA/ferrintei/rs6069753_case.raw")
colnames(gsa1) <- make.names(colnames(gsa1))

gsa_all <- rbind(gsa,gsa1)
gsa_all$X20.55025762_G <- round(gsa_all$X20.55025762_G)

```

```{r}
wes <- fread("~/WES/result/rs6069753_cnt.raw")
colnames(wes) <- make.names(colnames(wes))
wes <- wes %>% 
  select(IID,X20.55025762.G.A_G)
```


```{r}
target <- target4 %>% 
  select(IID.x, GeneticID,vcf_ID.y)


wes1 <- wes %>% 
  inner_join(target, by = c("IID" = "IID.x"))

```

```{r}
matome <- wes1 %>% 
  inner_join(gsa_all,by = c("vcf_ID.y" = "IID")) %>% 
  rename (WES = X20.55025762.G.A_G, GSA = X20.55025762_G ) %>% 
  mutate(WES = as.factor(WES)) %>% 
  mutate (GSA = as.factor(GSA))

```


```{r}
ggplot(matome, aes(x = WES, y = GSA)) + geom_jitter(width = 0.2,height = 0.2)
```

