---
title: "for_make_hailtarget"
author: "Takeo Naito,"
date: "6/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### read thiopurine phenotype file and make colnames better

```{r}
library(tidyverse)
library(readxl)
library(data.table)
thio <- read_xlsx("../data/Thiopurine toxicityforDalin_v2.xlsx")
colnames(thio) <- make.names(colnames(thio))

```
### delete duplicate IDs

```{r}
thio1 <- thio %>% 
  distinct(LAB.ID,.keep_all = TRUE)
```

### read WES data information (key for IID and genetic ID)

```{r}
wes <- read_xlsx("../data/niddk_vcf_samples_biostats.xlsx",
                 col_types = c("text"))
colnames(wes) <- make.names(colnames(wes))
```
### read Copy of Genetics 01_02_2019.xls file for information of All data

```{r}
pheno <- read_xls("../data/Copy of Genetics 01_02_2019.xls",col_types = "text") 
colnames(pheno) <- make.names(colnames(pheno))
pheno <- pheno %>% 
  drop_na(Genetic.ID)
dim(pheno)
```

### left join and make colnames names better 

```{r}
wes1 <- wes %>% 
  left_join(thio1,by = c("GeneticID" = "LAB.ID")) %>% 
  dplyr::rename("#IID" = Seq_ID) %>% 
  dplyr::select(-AKA,-TelAviv_Controls,-comment) %>% 
  dplyr::rename(Leukopenia = Leukopenia..Yes.1..No.0) %>% 
  dplyr::rename(GI.Intolerance = GI.Intolerance..N.V..abd.pain..diarrhea.) %>% 
  dplyr::rename(Allergy = Allergy..rash..high.fever..hypersensitivity.rxn.) %>% 
  dplyr::rename(Flu.like = Flu.like.sx..fever..malaise..nausea..HA.)

```

### replace u and U into NA
```{r}
wes1[ifelse(wes1 == "U",TRUE,FALSE)] <- -9
wes1[ifelse(wes1 == "u",TRUE,FALSE)] <- -9
```
### replace 0 and 1 into 1 and 2

```{r}
wes1[ifelse(wes1 == 1,TRUE,FALSE)] <- 2
wes1[ifelse(wes1 == 0,TRUE,FALSE)] <- 1
```
###  left_join with pheno

```{r}
wes2 <- wes1 %>% 
  left_join(pheno , by = c("GeneticID" = "Genetic.ID")) %>% 
  dplyr::select(-Genetic.Comments)

```
### confirm -9 value's row and column number

```{r}
which(wes2 == -9, arr.ind=T)
```
### replace -9 to NA

```{r}
wes3 <- wes2 %>% 
  replace(., . == -9, NA)

which(wes3 == -9, arr.ind=T)
```
### rename column name

```{r}
wes3 <- wes3 %>% 
  dplyr::rename(IID = "#IID")

colnames(wes3) <- str_replace_all(colnames(wes3), "\\.", "_")
```

# write targettxt
```{r}
write.table(wes3,"../data/targethail1.txt",sep = "\t", row.names = FALSE, quote = FALSE)
test <- fread("../data/targethail1.txt",fill = TRUE)
dim(test)
```




