---
title: "FOR_MAKE_Race_disease_cedars"
author: "Takeo Naito, MD PhD"
date: "7/26/2019"
output: html_document
---




```{r}
library(tidyverse)
library(readxl)
library(data.table)
library(readr)
library(knitr)
library(kableExtra)

```
```{r}
getwd()
```


# read WES data information (key for IID and genetic ID)

```{r}
wes <- read_xlsx("../data/niddk_vcf_samples_biostats.xlsx",
                 col_types = c("text"))
colnames(wes) <- make.names(colnames(wes))

dim(wes)
```
# left join and make colnames names better 

```{r}
wes1 <- wes %>% 
  dplyr::select(-AKA) 
dim(wes1)
```



# read Copy of Genetics 01_02_2019.xls file for information of All data for ethnicity

```{r}
pheno <- read_xls("../data/Copy of Genetics 01_02_2019.xls",col_types = "text") 
colnames(pheno) <- make.names(colnames(pheno))
pheno <- pheno %>% 
  distinct(Genetic.ID, .keep_all = TRUE)

dim(pheno)

```

# merge dataset (thiopurine and ethnicity)

```{r}
merge1 <- wes1 %>% 
  left_join(pheno, by = c("GeneticID" = "Genetic.ID"))

dim(merge1)

```
# read cleaned plink fam file

```{r}

fam <- fread("../data/cleaned_cedars.fam")

dim(fam)
```

# merge dataset again
```{r}
merge2 <- merge1 %>% 
  inner_join(fam,by = c("Seq_ID" = "V2"))

dim(merge2)
```






# Add diagnosis column(DIAG)
# Basially, priority of diagnosis is according to this order
  - 1. Current.Diagnosis
  - 2. Initial.IBD.Diagnosis
  - 3. DX
```{r}
merge3 <- merge2 %>% 
  mutate(DIAG = if_else(!is.na(Current.Diagnosis) & Current.Diagnosis != 'None',
                        Current.Diagnosis,
                        if_else(!is.na(Initial.IBD.Diagnosis), Initial.IBD.Diagnosis,
                                DX))) 
```


```{r}
merge4 <- merge3 %>% 
  select(Seq_ID,DIAG,Race,Relation.to.Proband,NOT.CEDARS,TelAviv_Controls,comment)

table(merge4$DIAG,useNA = "ifany")
```

```{r}
onlyibd <- merge4 %>% 
  filter(DIAG %in% c("Crohn's Colitis","Crohn's Disease","Ulcerative Colitis","Colitis Unclear Type/IBDU"))
# numbe of subjects
dim(onlyibd)

# number of diagnosis
table(onlyibd$DIAG)

# number of race
table(onlyibd$Race)

# number of relatedness
table(onlyibd$Relation.to.Proband)
```






# write tsv
```{r}
write_tsv(onlyibd,"/mnt/share6/FOR_Takeo/WES/acmg/onlyibd_subjects.txt")
```

