---
title: "SKAT_Target"
author: "Takeo Naito, MD PhD"
date: "7/20/2019"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(data.table)
library(readr)
library(knitr)
library(kableExtra)

```
```{r}
getwd()
```


# read thiopurine phenotype file and make colnames better
```{r}
thio <- read_xlsx("../data/Thiopurine toxicityforDalin_v2.xlsx")
colnames(thio) <- make.names(colnames(thio))
dim(thio)

thio1 <- thio %>% 
  distinct(LAB.ID,.keep_all = TRUE)
dim(thio1)

```
# read WES data information (key for IID and genetic ID)

```{r}
wes <- read_xlsx("../data/niddk_vcf_samples_biostats.xlsx",
                 col_types = c("text"))
colnames(wes) <- make.names(colnames(wes))

dim(wes)
```
# left join and make colnames names better 

```{r}
wes1 <- wes %>% 
  left_join(thio1,by = c("GeneticID" = "LAB.ID")) %>% 
  dplyr::select(-AKA,-TelAviv_Controls,-comment) %>% 
  dplyr::rename(Leukopenia = Leukopenia..Yes.1..No.0) %>% 
  dplyr::rename(GI.Intolerance = GI.Intolerance..N.V..abd.pain..diarrhea.) %>% 
  dplyr::rename(Allergy = Allergy..rash..high.fever..hypersensitivity.rxn.) %>% 
  dplyr::rename(Flu.like = Flu.like.sx..fever..malaise..nausea..HA.)

dim(wes1)
```
# replace u and U into NA
```{r}
wes1[ifelse(wes1 == "U",TRUE,FALSE)] <- -9
wes1[ifelse(wes1 == "u",TRUE,FALSE)] <- -9

```
# replace 0 and 1 into 1 and 2

```{r}
wes1[ifelse(wes1 == 1,TRUE,FALSE)] <- 2
wes1[ifelse(wes1 == 0,TRUE,FALSE)] <- 1
```


# read Copy of Genetics 01_02_2019.xls file for information of All data for ethnicity

```{r}
pheno <- read_xls("../data/Copy of Genetics 01_02_2019.xls",col_types = "text") 
colnames(pheno) <- make.names(colnames(pheno))
pheno <- pheno %>% 
  distinct(Genetic.ID, .keep_all = TRUE)

dim(pheno)

```

# merge dataset (thiopurine and ethnicity)

```{r}
merge1 <- wes1 %>% 
  left_join(pheno, by = c("GeneticID" = "Genetic.ID"))

dim(merge1)

```
# read cleaned plink fam file

```{r}

fam <- fread("../data/cleaned_cedars.fam")

dim(fam)
```

# merge dataset again
```{r}
merge2 <- merge1 %>% 
  inner_join(fam,by = c("Seq_ID" = "V2"))

dim(merge2)
```



# read cliical phenotype (Disease locatio and disease behavior)

```{r}
kyosaku <- read_tsv('/mnt/share6/FOR_Takeo/WES/cd_clean.txt')

```
# merge files and mutate FID column for SKAT 

```{r}
merge3 <- merge2 %>% 
  left_join(kyosaku,by = c('GeneticID' = 'genetic_id'))

merge3 <- merge3 %>% 
  mutate(FID = 0) %>% 
  rename(IID = Seq_ID) %>% 
  select(FID,IID,everything())
```

# Add diagnosis column(DIAG)
# Basially, priority of diagnosis is according to this order
  - 1. Current.Diagnosis
  - 2. Initial.IBD.Diagnosis
  - 3. DX
```{r}
merge3 <- merge3 %>% 
  mutate(DIAG = if_else(!is.na(Current.Diagnosis) & Current.Diagnosis != 'None',
                        Current.Diagnosis,
                        if_else(!is.na(Initial.IBD.Diagnosis), Initial.IBD.Diagnosis,
                                DX))) 
```
# Add column for complicated or not

```{r}
merge4 <- merge3 %>% 
  mutate(complicate = if_else(Disease.Classification.of.CD == 1, 1,2))
```
# select necessary columns

```{r}
cols <- colnames(merge4)

hoshi <- cols[c(1:15,51,52,36,37,27,49)]
hoshi

merge5 <- merge4 %>% 
  select(hoshi)
```

# make age columns as numeric
```{r}
merge5 <- merge5 %>% 
  mutate(Age.at.Onset = as.numeric(Age.at.Onset)) %>% 
  mutate(Exact.Age.at.Onset = as.numeric(Exact.Age.at.Onset))

```

# add columns about Disease location and Diagnosis for SKAT
```{r}

merge5 <- merge5 %>% 
  mutate(L1 = if_else(DzLocation == 1,2,1)) %>% 
  mutate(L2 = if_else(DzLocation == 2,2,1)) %>% 
  mutate(L3 = if_else(DzLocation == 3,2,1)) %>% 
  mutate(CDorHC = ifelse(DIAG == "Crohn's Disease",2,
                          ifelse(DIAG == "Non-IBD",1,NA))) %>% 
  mutate(UCorHC = ifelse(DIAG == "Ulcerative Colitis",2,
                          ifelse(DIAG == "Non-IBD",1,NA))) %>% 
  mutate(CDorUC = ifelse(DIAG == "Crohn's Disease",2,
                          ifelse(DIAG == "Ulcerative Colitis",1,NA))) %>% 
  mutate(Caucasian = ifelse(Race == "Caucasian",2,1)) %>% 
  select(-Race,-DIAG)
```

# make NA to -9 for SKAT

```{r}
merge5[ifelse(is.na(merge5),TRUE,FALSE)] <- -9
```


# delete umbigous sample
```{r}
merge5[3313,]
merge5 <- merge5 %>% 
  slice(-3313)
```



```{r}
merge5_bi <- merge5 %>% 
  select(-GeneticID,-Age.at.Onset,-Exact.Age.at.Onset)

merge5_con <- merge5 %>% 
  select(FID,IID,Age.at.Onset,Exact.Age.at.Onset)
  
```


# write table 

```{r}
write_tsv(merge5, "~/WES/data//SKAT/SKAT_target.txt")
write_tsv(merge5_bi, "~/WES/data//SKAT/SKAT_target_binary.txt")
write_tsv(merge5_con, "~/WES/data//SKAT/SKAT_target_coutinuous.txt")
```

