---
title: "subjects_who_have_no_phenotype"
author: "Takeo Naito, MD PhD"
date: "7/23/2019"
output: html_document
---



```{r}
library(tidyverse)
library(readxl)
library(data.table)
library(readr)
library(knitr)
library(kableExtra)

```
```{r}
getwd()
```


# read thiopurine phenotype file and make colnames better
```{r}
thio <- read_xlsx("../data/Thiopurine toxicityforDalin_v2.xlsx")
colnames(thio) <- make.names(colnames(thio))
dim(thio)

thio1 <- thio %>% 
  distinct(LAB.ID,.keep_all = TRUE)
dim(thio1)

```
# read WES data information (key for IID and genetic ID)

```{r}
wes <- read_xlsx("../data/niddk_vcf_samples_biostats.xlsx",
                 col_types = c("text"))
colnames(wes) <- make.names(colnames(wes))

dim(wes)
```
# left join and make colnames names better 

```{r}
wes1 <- wes %>% 
  left_join(thio1,by = c("GeneticID" = "LAB.ID")) %>% 
  dplyr::select(-AKA,-TelAviv_Controls,-comment) %>% 
  dplyr::rename(Leukopenia = Leukopenia..Yes.1..No.0) %>% 
  dplyr::rename(GI.Intolerance = GI.Intolerance..N.V..abd.pain..diarrhea.) %>% 
  dplyr::rename(Allergy = Allergy..rash..high.fever..hypersensitivity.rxn.) %>% 
  dplyr::rename(Flu.like = Flu.like.sx..fever..malaise..nausea..HA.)

dim(wes1)
```
# replace u and U into NA
```{r}
wes1[ifelse(wes1 == "U",TRUE,FALSE)] <- -9
wes1[ifelse(wes1 == "u",TRUE,FALSE)] <- -9

```
# replace 0 and 1 into 1 and 2

```{r}
wes1[ifelse(wes1 == 1,TRUE,FALSE)] <- 2
wes1[ifelse(wes1 == 0,TRUE,FALSE)] <- 1
```


# read Copy of Genetics 01_02_2019.xls file for information of All data for ethnicity

```{r}
pheno <- read_xls("../data/Copy of Genetics 01_02_2019.xls",col_types = "text") 
colnames(pheno) <- make.names(colnames(pheno))
pheno <- pheno %>% 
  distinct(Genetic.ID, .keep_all = TRUE)

dim(pheno)

```

# merge dataset (thiopurine and ethnicity)

```{r}
merge1 <- wes1 %>% 
  left_join(pheno, by = c("GeneticID" = "Genetic.ID"))

dim(merge1)

```
# read cleaned plink fam file

```{r}

fam <- fread("../data/cleaned_cedars.fam")

dim(fam)
```

# merge dataset again
```{r}
merge2 <- merge1 %>% 
  inner_join(fam,by = c("Seq_ID" = "V2"))

dim(merge2)
```



# read cliical phenotype (Disease locatio and disease behavior)

```{r}
kyosaku <- read_tsv('/mnt/share6/FOR_Takeo/WES/cd_clean.txt')

```
# merge files and mutate FID column for SKAT 

```{r}
merge3 <- merge2 %>% 
  left_join(kyosaku,by = c('GeneticID' = 'genetic_id'))

merge3 <- merge3 %>% 
  mutate(FID = 0) %>% 
  rename(IID = Seq_ID) %>% 
  select(FID,IID,everything())
```

# Add diagnosis column(DIAG)
# Basially, priority of diagnosis is according to this order
  - 1. Current.Diagnosis
  - 2. Initial.IBD.Diagnosis
  - 3. DX
```{r}
merge3 <- merge3 %>% 
  mutate(DIAG = if_else(!is.na(Current.Diagnosis) & Current.Diagnosis != 'None',
                        Current.Diagnosis,
                        if_else(!is.na(Initial.IBD.Diagnosis), Initial.IBD.Diagnosis,
                                DX))) 
```
# Add column for complicated or not

```{r}
merge4 <- merge3 %>% 
  mutate(complicate = if_else(Disease.Classification.of.CD == 1, 1,2))
```
# select CD subjects who have no phenotype

```{r}
df1 <- merge4 %>% 
  filter(DIAG == "Crohn's Disease") %>% 
  filter(is.na(Disease.Location)) %>% 
  select(FID,IID,GeneticID,Disease.Location) %>% 
  select(GeneticID)



df2 <- merge4 %>% 
  filter(DIAG == "Crohn's Disease") %>% 
  filter(is.na(Disease.Classification.of.CD)) %>% 
  select(FID,IID,GeneticID,Disease.Classification.of.CD) %>% 
  select(GeneticID)

df3 <- merge4 %>% 
  filter(DIAG == "Crohn's Disease") %>% 
  filter(is.na(DzLocation)) %>% 
  select(FID,IID,GeneticID,DzLocation) %>% 
  select(GeneticID)

df4 <- merge4 %>% 
  filter(DIAG == "Crohn's Disease") %>% 
  filter(is.na(DzBehavior)) %>% 
  select(FID,IID,GeneticID,DzBehavior) %>% 
  select(GeneticID)
```
# do full join 

```{r}
dftotal <- df1 %>% 
  full_join(df2,by =c("GeneticID")) %>% 
  full_join(df3,by =c("GeneticID")) %>% 
  full_join(df4, by =c("GeneticID"))

```
# write tsv
```{r}
write_tsv(dftotal,"/mnt/share6/FOR_Takeo/WES/VEOIBD/unknown_phenotype_subjects.txt")
```

